---
title: "StringR Basics"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```


## Counting string {.build}
```{r setup_count, include=FALSE}
df_input2 %>% count(health_status, sort = TRUE) %>% slice(2) %>% select(1) %>% pull()
#restart here
```

`str_count()` is a function we can use to count the number of rows that match a particular pattern.  
  
The output with either be 1 (match), or 0 (no match)  

<div class= "red2"> Example </div>  
In this code:  
- string we want to evaluate is ``r dataframe_joined_name`$`r df_joined_string3_name``  
- pattern we want to count `"High Cholesterol"`  

`str_count(heart_joined$health_status, "High Cholesterol")`

## Counting
`str_count(heart_joined$health_status, "High Cholesterol")`

```{r str_count example, echo= FALSE, warning = FALSE}
str_count(heart_joined$health_status, "High Cholesterol")
```

## Summarizing our counts
A bunch of 0 and 1 are not incredibly useful.

But since R is good at adding, we can simply wrap the previous expression in `sum()`  

Try it below:  
```{r sum_string, exercise = TRUE}

```

## Matching subsets of strings {.build}

We previously matched the entire string "High Cholesterol"  
But we can use the same function to detect patterns within longer strings.  
  
Let's look for how many patients take a statin of any kind using
`str_count(heart_joined$medication_hx, "statin")` 

## What about subsets of strings?  
`str_count(heart_joined$medication_hx, "statin")` 

What does the ouput mean?  
```{r str_count_exercise, exercise = TRUE}

```

## A note about string patterns

When using a stringr function, you may get an output saying a string pattern doesn't exist. If you know for sure it does, <div class="blue"> double check capitalization </div>.  
  
**The string must match exactly, or it will not be found!**

## stringr Exercise
How many people having an `"auntie" or "aunt"` in their health history?  
  
Go to code/  
Open 09_stringr.Rmd  
Complete the exercise. 



## Detecting strings {.build}
In addition to counting, we can use another function `str_detect()` to logically evaluate a character string.  
  
Because this logically evaluates an expression, the output is either `TRUE or FALSE`  
  
Practially, `str_detect` is used to detect the presence or absence of a pattern in a string  

## Logic Evaluation 
<div class= "red2"> Example </div>  
Find the patients with diabetes using the following code  
  
`str_detect(heart_joined$health_status, "Diabetic")`  
  
```{r strdetect, exercise=TRUE}

```

## Modifiying strings with `str_replace()` {.build}
In the health_status column we have:  
-`"Diabetic"`  
-`"High Cholesterol"`  
-`"Normal blood sugar and cholesterol"`  
  
But let's say we want to simplify healthy individuals to `"Normal"`

`str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "Normal")`

```{r str_replace, exercise=TRUE}

```

## Modifiying strings with `str_replace()` {.build}
We use this same code to  modify the `health_status` column by assigning it to the same variable     
  
`heart_joined$health_status <-`  
       `str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "normal")`

```{r str_replace in col, warning = FALSE, echo=FALSE}
heart_joined$health_status <- 
  str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "normal")

head(heart_joined$health_status, n = 10)
```

## Using `stringr` with `dplyr` {.build}
We can use `stringr` functions in tandem with `dplyr` functions.  
  
<div class= "red2"> Example </div>  
We want to make a logical variable (`TRUE`/`FALSE`) that tells us if a patient has a normal health history using  

`heart_joined2 <- heart_joined %>% mutate(healthy = str_detect(health_status, "normal"))` 

```{r stringr with dplyr, echo=FALSE, warning=FALSE}
heart_joined2 <- mutate(heart_joined, healthy = str_detect(health_status, "normal"))

head(heart_joined2$healthy, n = 10)
```
 