---
title: "Tidybiology Tiny Lectures: ggplot "
subtitle: "ggplot basics"
author: "Matthew Hirschey, Ph.D."
date: "April 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

# Welcome back

# "The simple graph has brought more information to the data analyst’s mind than any other device."  -- John Tukey

## Visualize Data with
```{r ggplot_logo, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", out.height = '100%'}
url <- "https://github.com/tidyverse/ggplot2/raw/master/man/figures/logo.png"
knitr::include_graphics(url)
```
<font size = "1">[https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org)</font>

## A basic plot using `ggplot` in R
We ended the last class with:

```{r ggplot_basic, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ggplot(heart, aes(x = sex, y = chol)) +
  geom_point(color = "darkblue", size = 3) +
  labs(x = "sex", 
       y = "Cholesterol", 
       title = "Cholesterol values from the heart disease dataset", 
       caption = "Data from Kaggle | Plot from @matthewhirschey") +
  theme_minimal() +
  NULL
```

## which built this...
```{r ggplot_basic2, message=FALSE, warning=FALSE, echo=FALSE}
#need to parameterize this; embed rmd here?; but still need to parameterize that doc.
ggplot(heart, aes(x = sex, y = chol)) +
  geom_point(color = "darkblue", size = 3) +
  labs(x = "sex", y = "Cholesterol", title = "Cholesterol values from the heart disease dataset", caption = "Data from Kaggle | Plot from @matthewhirschey") +
  theme_minimal() +
  NULL
```

# Basic ggplot2
Any questions from last week?

## Basics of a ggplot code {.build}

Below is an example of the most basic form of the ggplot code  
  
`ggplot(data = dataframe) +`   
  `geom(mapping = aes(x, y))`  
  
Take a moment to look back at the code template.  You can see that in that code we assigned a `dataset` and the information we needed to `map` it to a type of plot

## Basics of a ggplot code {.build}
  
`ggplot(data = `r dataframe_name`) +`  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  
  
```{r simple_plot, echo = FALSE, warning = FALSE}
df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 
```

## Build a `ggplot` {.build}
- "Initialize" a plot with `ggplot()`  
  
`ggplot(data = `r dataframe_name`) +`  
  
- Add layers with `geom_functions`  
  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  

## Notes on `ggplot` style {.build}
- Generally, different people have **strong** opinions about style and data visualization  
  
- Data visualization is a rich and complex area of study and is beyond the scope of this introductory course  
  
**That being said, here are a few style tips:**    
- While you can put the `+` at the beginning of the next line, it is generally put at the end of the previous line    
  
- Arguments do not need to be explicit: this chunk of code is equivalent to the previous chunk (w/ data, mappings)  
`ggplot(`r dataframe_name`) +`    
`geom_point(aes(x = `r df_numeric2_name`, y = `r df_numeric2_name`))`  

