---
title: "Tidybiology Tiny Lectures: ggplot "
subtitle: "ggplot: global vs local"
author: "Matthew Hirschey, Ph.D."
date: "April 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```


## Global vs Local
```{r global_v_local_concepts, echo = FALSE, warning = FALSE}
ggplot(df_input) +
  geom_jitter(aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_boxplot(aes(!!sym(df_char1_name), !!sym(df_numeric1_name)))
```

## Global vs Local
However, if we map our x and y values in the ggplot function we find that we generate the same graph  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot()`  
```{r global_v_local, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot()
```

## Global vs Local {.build}
This is because when you set the aes mappings in the original `ggplot` function you are setting the `aes` **globally**.  
This means all the functions afterwards will inherit that mapping. So in our example, this means that both the jitter and boxplot geoms know to graph the same information  
  
You can also set aes values **locally** within the geom function. Doing so will only change the values in that geom

## Global vs Local
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`))`      
```{r global_v_local1, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name)))
```

## Global vs Local {.build}
```{r global_v_local2, echo=FALSE}
mean <- mean(df_numeric2_vec)
sd <- sd(df_numeric2_vec)
```

Data can also be set locally or globally. For this example, let's filter our original data first using the `dplyr::filter` function
  
`df_filter <- `r dataframe_name` %>% filter(`r df_numeric2_name` > `r round(mean + 2*sd)`)`  
  
*this number is two standard deviations above the `mean` value of `r df_numeric2_name`

## Global vs Local {.build}
Now, let's identify only the `r dataframe_about` in our data that are outliers, more than 2SD above the mean, by setting data locally in a new geom  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`   
`geom_label(data=df_filter, aes(label=`r df_id_name`))`  

## Global vs Local {.build}
```{r global_v_local3, echo = FALSE, warning = FALSE}
df_filter <- df_input %>% filter(!!sym(df_numeric2_name) > round(mean + 2*sd))

ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  geom_label_repel(data = df_filter, aes(label = !!sym(df_id_name)))
```

## Global vs Local {.build}
You notice we have to indicate the new dataset, but because it has the same x and y values, we did not need to set those mappings  
  
Go to code/  
Open 06_global_v_local.Rmd  
Complete the exercise to practice mapping locally and globally.   

