---
title: "Tidybiology Tiny Lectures: ggplot "
subtitle: "ggplot: themes"
author: "Matthew Hirschey, Ph.D."
date: "April 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Themes {.build}
You can change almost everything you see on your chart, but a lot of the things you may look to change are part of the "theme"  
  
Here we are going to change some features about our title text:  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12))` 

## Themes {.build}
```{r themes, echo = FALSE, warning = FALSE}
theme1 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12)) 
original_plot + theme1
```

## Themes {.build}
Next, let's change the aesthetics of our legend box  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"))`   

## Themes {.build}
```{r themes2, echo = FALSE, warning = FALSE}
theme2 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black")
  ) 
theme1 + theme2
```

## Themes {.build}
Finally, let's change the legend position  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"),`  
`legend.position = "bottom"))` 

## Themes {.build}
```{r themes3, echo = FALSE, warning = FALSE}
theme3 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black"),
        legend.position = "bottom"
  ) 
theme2 + theme3
```

## Themes {.build}
Pre-set themes also exist as an easy way to change the entire theme of your graph quickly. They can also be combined with custom theme settings  

`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme_minimal()`  

## Themes {.build}
```{r themes4, echo = FALSE, warning = FALSE}
theme4 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme_minimal()
theme1 + theme4
```



