---
title: "Introduction to Tidyverse"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```
## Tidyverse 
Tidyverse is one suite of tools for data science  
  
<center><iframe src="https://www.tidyverse.org"></iframe></center>

## Exploratory Data Analysis
```{r r4ds, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/hadley/r4ds/raw/master/diagrams/data-science.png"
knitr::include_graphics(url)
```

## Tidyverse Basic Principles
**IMPORT** (readr):  
- `read_csv()`  
- `read_delim()`  
  
**TIDY & TRANSFORM** (dplyr):  
- `mutate()` adds new variables that are functions of existing variables  
- `select()` picks variables based on their names.  
- `filter()` picks cases based on their values.  
- `summarise()` reduces multiple values down to a single summary.  
- `arrange()` changes the ordering of the rows.  
  
**VISUALISE** (ggplot): creating graphics, based on ‘The Grammar of Graphics’  
- `aes()`  
- `geom_x()` + layers   
  
**MODEL** (broom):   
- `tidy()`, `glance()`, `augment()`  

----
```{r pipe, echo=FALSE, fig.cap="", out.width='77%'}
url <- "https://s3-us-west-2.amazonaws.com/collections.lacma.org-images/remote_images/ma-150089-WEB.jpg"
knitr::include_graphics(url)
```

## `magrittr` package
```{r magrittr, echo=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/tidyverse/magrittr/raw/master/man/figures/logo.png"
knitr::include_graphics(url)
```

## `magrittr` package
`magrittr` package by Stefan Milton Bache developed the concept of the pipe, which is used heavily in the `tidyverse`  
  
  
```{r magrittr2, echo=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/tidyverse/magrittr/raw/master/vignettes/magrittr.jpg"
knitr::include_graphics(url)
```
<br>  
<center><font size = "10">"and then"</font></center>

## The Pipe {.build}
  The “pipe” is a sequence of functions, that are sequentially applied to an object  
```{r wakeup, echo=TRUE, eval=FALSE}
wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  go(to = "work")
```
  
  Alternative nested code:  
```{r wakeup_nested, echo=TRUE, eval=FALSE}
go(eat(put_on(wakeup(self), "clothes"), "breakfast"), to = "work")
```

## The Pipe (Quiz I)
What does this code do? 
  
```{r wakeup_quiz, echo=TRUE, eval=FALSE}
wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  fmk() %>% 
  go(to = "work")
```
  
## The Pipe (Quiz II)
What does this code do? 
  
```{r wakeup_quiz2, echo=TRUE, eval=FALSE}
morning_routine <- wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  fmk() %>% 
  go(to = "work")
```

## The Pipe (Quiz III)
What does this code do? 

``r dataframe_name` %>%`  
`select(`r colnames(df_input)[2]`, `r colnames(df_input)[3]`) %>%`  
`group_by(`r colnames(df_input)[3]`) %>%`  
`summarize(mean(`r colnames(df_input)[2]`))`  
  
Try it out here!
```{r pipe_ex, exercise=TRUE}

```
## Writing code IS NOT like drawing an owl
```{r owl, echo=FALSE, fig.cap="", out.width='50%'}
url <- "https://66.media.tumblr.com/tumblr_l7iwzq98rU1qa1c9eo1_500.jpg"
knitr::include_graphics(url)
```

## Writing code IS a step-wise process
```{r lego, echo=FALSE, fig.cap="", out.width='70%'}
#change this to a good "and then" example
url <- "https://github.com/rstudio-education/datascience-box/raw/master/slides/u1_d01-meet-the-toolkit/img/lego-steps.png"
knitr::include_graphics(url)
```
<font size = "1">https://datasciencebox.org</font>
