---
title: "Introduction to R"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

# R <br> language
Any questions?

## R is a language
```{r language, echo=FALSE, fig.cap=""}
url <- "https://source.unsplash.com/ZzWsHbu2y80/256x455"
knitr::include_graphics(url)
```
<font size = "1">Photo by Hannah Wright on Unsplash</font>

## R has values
- 1
- "North Carolina"
- "2020-04-08"

## R has objects
- A name without quotes  
- Assigned using `<-` (looks like an arrow pointing left)  
- Can be a value, object, or function result  

## Try assigning an object 
1. Assign an object; **remember, no quotes on name**  
`name <- 4`  
  
2. Return that object by typing its name  
`name`  
  
Try this in the code chunk below, then hit "Run Code"  

```{r my_first_object, exercise=TRUE}

```

## R has functions
- A name without quotes  
- Followed by () to run the function  
- Optional arguements: values, objects, or function results  
- `round(x, digits = 3)`

Try this in the code chunk below, then hit "Run Code"  

```{r my_first_fun, exercise=TRUE}
round(pi, digits = 3)

```

## Which of these are numbers?

```{r num_quiz, echo=FALSE}
question("Which of these are numbers?",
  answer("1", correct = TRUE),
  answer('"1"', message = "Because it has quotes, it is an string"),
  answer('"one"', message = "Because it has quotes, it is an string"),
  answer("one",  message = "R recognizes this as an object"), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```

## Which of these will work?
Suppose `one <- 1`
```{r one_quiz, echo=FALSE}
question("Which of these will work?",
  answer("log(1)", correct = TRUE),
  answer('log("1")', message = "You cannot log transform a word (string)"),
  answer('log("one")', message = "You cannot log transform a word (string)"),
  answer("log(one)", correct = TRUE), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```
Try it for yourself!  
  
```{r scratchpad, exercise=TRUE}
one <- 1
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r kable}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  knitr::kable()
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r tibble}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  print(as_tibble())
```

## Extract or create new objects
```{r extract_setup, include=FALSE}
col <- sample(colnames(df_input), 1)
```

You can call a single part of the data frame  
  
`r dataframe_name`$`r col`

```{r extract, echo = FALSE, warning = FALSE}
df_input %>% 
  select(!!col)
```

## Extract or create new objects 
  
Write the R code required to extract a variable from the `r dataframe_name` dataset:  
  
Remember, the format is: `r dataframe_name`$`r col`  
  
```{r extract_exercise, exercise=TRUE}

```

## Extract or create new objects 
  
You can also save a part of the dataframe as an object for later use  
  
``r col` <- `r dataframe_name`$`r col``  
  
In the code chunk below:  
1. On the first line, write the R code to save a single column to a new object   
2. On the second line, type the object name - this will print out the new object  
3. Run the code  
  
```{r create_exercise, exercise=TRUE}

```