---
title: "Saving Thing in R"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Saving a new dataset as an object

If you'd like to save the output of your wrangling as an object, you will need to use the `<-` operator

``r dataframe_name`_new <- `r dataframe_name` %>%`  
            `group_by(`r colnames(df_character_select)`) %>%`  
            `summarise(mean(`r colnames(df_numeric_select)`))`  
  
*Assigning the object the same name (i.e. ``r dataframe_name``) will overwrite the object  
  
```{r df_object_exercise, exercise=TRUE}

```
  
## Saving a new dataset as a file

To save ``r dataframe_name`_new` as a new file (e.g. csv), run the following:  
  
`write_csv(`r dataframe_name`_new, "`r dataframe_name`_new.csv")`
  