---
title: "Introduction to Mutate"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## `mutate()` {.build}

The `mutate()` verb, unlike the ones covered so far, creates new variable(s) i.e. new column(s). For example

``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(1)`  
  
The code chunk above takes all the elements of the column `r colnames(df_numeric_select)`, evaluates the square root of each element, and populates a new column called `new_col` with these results
  
## Try `mutate()` to make a new column
  
``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(3)`  
  
```{r mutate_exercise, exercise=TRUE}

```
