---
title: "Assessing Cholesterol Levels in the Heart Study"
subtitle: "Rmarkdown Parameter Exercise"
output: html_document  
params: 
  age: 40
  gender: "male"
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidybiology)
library(viridis)

#clear environment
#rm(list=ls()) 
```

```{r import_data, include=FALSE, message=FALSE, warning=FALSE}
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

```{r calculations, include=FALSE, message=FALSE, warning=FALSE}
##### START HERE #####

#What percent of patients in this dataset are male v. female?
heart %>% 
  count(sex, sort = TRUE)

freq <- heart %>% 
  group_by(sex) %>% 
  summarize(n = n()) %>% 
  mutate(freq = n/sum(n))

#What is the median age of patients
median <- median(heart$age)

#What is the average cholesterol level of patients at the median age?
avg_chol <- heart %>% 
  filter(age == median(heart$age)) %>% 
  summarize(mean_cholesterol = mean(chol))

#Store a new dataframe object that groups patients by sex and age, and then calculates the average cholesterol levels for patients in this group
age_groups <- heart %>% 
  group_by(sex, age) %>% 
  summarize(mean_cholesterol = mean(chol))

my_age_chol <- heart %>% 
  filter(between(age, params$age-2, params$age+2), 
         sex == params$gender) %>% 
  summarize(mean_cholesterol = mean(chol))

older <- if_else(
  params$age > median, "older", "younger"
)

#store some of these values, so that you can call them inline below
```

In the heart study data set, the proportion of females is `r round(freq %>% filter(sex == "female") %>% select(freq) %>% pull(), 1)` and males is `r round(freq %>% filter(sex == "male") %>% select(freq) %>% pull(), 1)`. The median age of patients in this data set is `r median`, which have an average serum cholesterol level of `r round(avg_chol, 0)` mg/dl. I self-identify as a `r params$age` year old `r params$gender`, which is `r older` than the median age of patients in this data set. The average cholesterol for patients in this data set of someone my age is `r round(my_age_chol, 1)` mg/dl. 

```{r provenance, include=FALSE}
#Session information for provenance and reproducibility
session_provenance()
```
