---
title: "Introduction to Filter"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## `filter()` rows based on certain condition(s) {.build}

The `filter()` verb evalutes a logical statement, and if a row meets the condition of this statement (i.e. is true) then it gets chosen (or "filtered").  
  
  
**All other rows are discarded**  

## `filter()` {.build}

Filtering can be performed on **categorical** data

``r dataframe_name` %>%`     
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`     
    `head(3)`  

```{r filter_exercise, exercise=TRUE}

```

Note that `filter()` only applies to rows, and has no effect on columns

## `filter()` {.build}

Filtering can also be performed on **numeric** data

For example, if you wanted to choose `r colnames(df_numeric_select)` with a value greater than `r df_numeric_filter`, you would run the following. 

``r dataframe_name` %>%`     
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`     
    `head(3)`  

```{r filter_exercise2, exercise=TRUE}

```

## `filter()` {.build}

To filter on multiple conditions, you can write a sequence of `filter()` commands

``r dataframe_name` %>%`       
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`  
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`    
    `head(3)`  

```{r filter_exercise3, exercise=TRUE}

```

## `filter()` {.build}

To avoid writing multiple `filter()` commands, multiple logical statements can be put inside a single `filter()` command, separated by commas

``r dataframe_name` %>%`           
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`",`  
           ` `r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`      
    `head(3)`  

```{r filter_exercise4, exercise=TRUE}

```
