---
title: "Getting to know your data"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

# R <br>Getting to know your data
Any questions?

## ``r dataframe_name`` dataset
**`r dataframe_title`**  
  
- For this class, we will use a dataset called ``r dataframe_name``.  
  
- `r dataframe_source`  
  
- `r dataframe_desc`  
  
## Inspecting your dataframe: dimensions

Use the `dim()` function to see how many rows (observations) and columns (variables) are in ``r dataframe_name``

`dim(`r dataframe_name`)` 
  
Enter your function here:  
```{r dim_exercise, exercise=TRUE}

```

## Inspecting your dataframe: glimpse

Use the `glimpse()` function to see what kinds of variables the ``r dataframe_name`` dataset contains   
  
`glimpse(`r dataframe_name`)`  
  
Enter your function here:  
```{r glimpse_exercise, exercise=TRUE}

```
  
## Basic Data Types in R {.build}

Recall that R has different data types -

**character** - `"a"`, `"tidyverse"`  

**numeric** - `2`, `11.5`  

**integer** - `2L` (the `L` tells R to store this as an integer)  

**logical** - `TRUE`, `FALSE`  

**complex** - `1+4i`  

(**raw**)

You will also come across the **double** datatype. It is the same as **numeric**

**factor**. A **factor** is a collection of *ordered* character variables

## Basic Data Types in R

In addition to the `glimpse()` function, you can use the `class()` function to determine the data type of a specific column

`class(`r dataframe_name`$`r colnames(df_character_select)`)`

```{r class, echo = FALSE, warning = FALSE}
class(df_character_select[[1]])
```
  
Try getting the `class` of a variable:  
```{r class_exercise, exercise=TRUE}

```
