---
title: "Introduction to Joins"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Combining datasets {.build}

There are many times when you have two or more overlapping datasets that you would like to combine

The `dplyr` package has a number of `*_join` functions for this purpose

```{r left_join, echo=FALSE, fig.cap="", out.height='50%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```


## `left_join` {.build}

Returns all rows from a, and all columns from a and b

Rows in a with no match in b will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned

## `left_join` example {.build}

Take a look at the variables in each dataset - `r dataframe_name` and `r dataframe_join_name`  

You will notice that both datasets contain common variable - `r df_id_name`. This can therefore serve as a common variable to join on. Let's join on this:  
  
`left_join` `r dataframe_name` with `r dataframe_join_name` and assign the output to a new object called ``r dataframe_name`_join_left`

Go to code/  
Open 08_import_and_join.Rmd  
Complete the exercise to join the two datasets.  
  
Now you have one dataset with additional useful information

## `right_join`
```{r right_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```


## `right_join` {.build}
Returns all rows from b, and all columns from a and b

Rows in b with no match in a will have NA values in the new columns

If there are multiple matches between a and b, all combinations of the matches are returned
  
This is conceptually equivalent to a `left_join`, but can be useful when stringing together multiple steps using %>% 

## `inner_join`
```{r inner_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```

## `inner_join` {.build}
Returns all rows from a where there are matching values in b, and all columns from a and b

If there are multiple matches between a and b, all combination of the matches are returned

## `full_join`
```{r full_join, echo=FALSE, fig.cap="", out.height='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/dplyr-joins.png"
knitr::include_graphics(url)
```

## `full_join` {.build}
Returns all rows and all columns from both a and b  

Where there are no matching values, returns NA for the one missing  