---
title: "Introduction to Arrange"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## `arrange()` sorts rows {.build}

The input for arrange is one or many columns, and `arrange()` sorts the rows in ascending order i.e. from smallest to largest

For example, to sort rows from smallest to largest `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`   
  `arrange(`r colnames(df_numeric_select)`) %>%`   
  `head(3)`  

```{r arrange_exercise, exercise=TRUE}

```

## `arrange()` {.build}

To reverse this order, use the `desc()` function within `arrange()`

``r dataframe_name` %>%`  
  `arrange(desc(`r colnames(df_numeric_select)`)) %>%`  
  `head(3)`  

```{r arrange_exercise2, exercise=TRUE}

```
