---
title: "Counting COVID"
subtitle: "Key elements of the analytical toolbox for understanding COVID-19"
author: "Matthew Hirschey and Akshay Bareja"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## COVID-19 rapidly infected people across the US
```{r covid, echo=FALSE, fig.cap="", out.height = '100%'}
#covid map from AKSHAY
```

## Rapid emergence generated several healthcare decision-making challenges {.build}
**Volume**  
- Rapid data collection & storage allowed access to huge amounts of medical information  
  
**Ubiquity**  
- Data became available anywhere across geography, social, and economic classes   
  
**Latency**  
- Technology facilitates no delay in access to data    
  
## How do you make sense of COVID-19 data to make better health care decisions? {.build}
**Data-driven** decision making!  
  
```{r hierarchy, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>
  