---
title: "Introduction to Data Science"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Doctors make decisions based on symptoms 
```{r intro, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/701-FJcjLAQ/500x500"
knitr::include_graphics(url)
```
<font size = "1">Photo by [National Cancer Institute on Unsplash](https://unsplash.com/@nci)</font>

## New digital healthcare era introduces new decision-making challenges {.build}
**Volume**  
- Data collection & storage allows access to huge amounts of medical information  
  
**Ubiquity**  
- Data are available anywhere across geography, social, and economic classes   
  
**Latency**  
- Technology facilitates no delay in access to data    
  
## How do you make better health care decisions? {.build}
**Data-driven** decision making!  
  
```{r hierarchy, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>
  