---
title: "Introduction to Select"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## `select()` specific columns from your dataset {.build}
The most basic `select()` is one where you comma separate a list of columns you want included

For example, if you only want to select the `r colnames(df_character_select)` and `r colnames(df_numeric_select)` columns, run the following code chunk

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`  
    `head()`
    
```{r select_exercise, exercise=TRUE}

```

## `select()` {.build}

If you want to select all columns *except* `r colnames(df_character_select)`, run the following

``r dataframe_name` %>%`  
    `select(-`r colnames(df_character_select)`) %>%`  
    `head(5)`

```{r selectnot_exercise, exercise=TRUE}

```

## `select()` {.build}

Finally, you can provide a range of columns to return two columns and everything in between. For example

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`:`r colnames(df_numeric_select)`) %>%`  
    `head(5)`
    
```{r selectrange_exercise, exercise=TRUE}

```
