---
title: "ggplot practice"
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
#Load datasets we will use for the lecture
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

# Lets take a glimpse at our dataset called `heart`
```{r}

```

# Basics of a ggplot code
Below is an example of the most basic form of the ggplot code
```{r, eval=FALSE}
ggplot(data)+ #this will generate a blank plot for your data. At this point, ggplot knows what to plot, but not how.
  geom(mapping=aes(x, y)) #adding a geom tells ggplot how you would like to map your data
```

Take a moment to use this template to make a simple ggplot. The data `heart` is defined, but you can choose the variables you want to map to x and y. I would recommend using `geom_point`.  
  
Put your code in the code chunk here, run it by clicking the green arrow.  
```{r}


```

# Aesthetics
Everything up to this point gets you a basic graph; but what about colors, shapes and overall style?  
  
You can change 5 basic aesthetics  
1. Color- changes the outline color of your datapoints  
2. Size - choose the size of the datapoint  
3. Shape - choose a pre-defined shape  
4. Alpha- changes the transparency of each point  
5. Fill- changes the fill color of your points  

## Color- changes the outline color of your datapoints  
Retype or paste the exact code from the last code chunk, then add a color aesthetic. In the template, it would go here:  
`geom(mapping=aes(x, y), color = "my_favorite_color")`
```{r}

```

## Size - choose the size of the datapoint  
Retype or paste the exact code from the last code chunk, then add a size aesthetic. In the template, it would go in the same place as color. 
```{r}

```

## Shape - choose a pre-defined shape  
Retype or paste the exact code from the last code chunk, then add a shape aesthetic. In the help tab, type "shape" to learn more about the shape options available, and how to call them. 
You can also look here: https://ggplot2.tidyverse.org/reference/aes_linetype_size_shape.html
```{r}

```

## Alpha- changes the transparency of each point  
Retype or paste the exact code from the last code chunk, then add an alpha aesthetic. In the template, it would go in the same place as color. Alpha is a scale from 0 to 1, similar to a percentage scale.
```{r}

```

## Fill- changes the fill color of your points  
Retype or paste the exact code from the last code chunk, then added a fill color aesthetic. In the template, it would go in the same place as color. How is fill different than color?
```{r}

```
