---
title: "aes practice"
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
```

```{r, include=FALSE}
#Load datasets we will use for the lecture
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

# Lets take a glimpse at our dataset called `heart`
```{r}

```

# Basics of a ggplot code
Below is an example of the most basic form of the ggplot code
```{r, eval=FALSE}
ggplot(data)+ #this will generate a blank plot for your data. At this point, ggplot knows what to plot, but not how.
  geom(mapping=aes(x, y)) #adding a geom tells ggplot how you would like to map your data
```

Take a moment to use this template to make a simple ggplot. The data `heart` is defined, but you can choose the variables you want to map to x and y. I would recommend using `geom_point`.  
  
Put your code in the code chunk here, run it by clicking the green arrow.  
```{r}


```
  
Compare the code for these two plots
```{r}
plot1 <- ggplot(heart) +
  geom_point(aes(x = age, y = chol), color = "blue")

plot2 <- ggplot(heart) +
  geom_point(aes(x = age, y = chol, color = sex))

plot1 + plot2
```

Of the 5 basic aesthetics, 4 can be mapped to variables in your data
1. Color- changes the outline color of your datapoints  
2. Size - choose the size of the datapoint  
3. Shape - choose a pre-defined shape  
4. Alpha- cCANNOT BE MAPPED TO A VARIABLE. 
5. Fill- changes the fill color of your points  
  
In this code chunk, take either your plot from above, or the example I provided, and map your variables to the aesthetics. Make sure you map the variable *inside* the `aes()`. 
```{r}

```
  
Curious for more? Map more than two variables to two distinct aesthetics
```{r}

```


