---
title: "Global v. Local practice"
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
#Load datasets we will use for the lecture
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

# Lets take a glimpse at our dataset called `heart`
```{r}

```

# Basics of a ggplot code
Below is an example of the most basic form of the ggplot code
```{r, eval=FALSE}
ggplot(data)+ #this will generate a blank plot for your data. At this point, ggplot knows what to plot, but not how.
  geom(mapping=aes(x, y)) #adding a geom tells ggplot how you would like to map your data
```

Take a moment to use this template to make a simple ggplot. The data `heart` is defined, but you can choose the variables you want to map to x and y. Add a second geom. 
  
Put your code in the code chunk here, with LOCAL variables for each geom; run it by clicking the green arrow.  
```{r}


```
  
Put your code in the code chunk here, with GLOBAL variables for each geom; run it by clicking the green arrow.  
```{r}

```

The `geoms` are hierarchical, so they will inhert the data from the step above, unless you specficy differently. 
