---
title: "ggplot: putting it all together"
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
#Load datasets we will use for the lecture
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

# Lets take a glimpse at our dataset called `heart`
```{r}

```

# Basics of a ggplot code
Below is an example of the most basic form of the ggplot code
```{r, eval=FALSE}
ggplot(data)+ #this will generate a blank plot for your data. At this point, ggplot knows what to plot, but not how.
  geom(mapping=aes(x, y)) #adding a geom tells ggplot how you would like to map your data
```

Use this template to make a simple ggplot.  
  
1. Use the `heart` dataset, 
2. Make a scatterplot comparing age vs. maximum heart rate
Recall that a "data dictionary" is located here: https://github.com/matthewhirschey/tidybiology-plusds

```{r first_plot}

```

Next, fit a smooth line (linear regression) to these points
*by adding another `geom` and specifying the argument `method = "lm"` 
```{r}

```
  
Next, color by whether the patient has exercise-induced angina
```{r}

```

The last plot probably has your color based on a continuous scale; now change color by sex (a discrete variable)
```{r}

```

Next, fix the labels to capitalize, and be more informative
```{r}

```

Next, give the plot a title, by calling `title = "title"` within `labs`
And give the plot a caption, by calling `caption = "Plot by YOURNAME"` within `labs`
```{r}

```

Next, use `facet_grid` to split this graph based on whether the pateint has exercise induced angina or not. Make sure to wrap your variable in `factor()` to convert the numeric into a factor variable. 
```{r}

```

Next, change the theme to `light`
```{r}

```

Lastly, save the object as a png image using `ggsave`
```{r}

```
