---
title: "Tidybiology +DS: Session 2 "
subtitle: "Introduction to Data Science in Health Care using R"
author: "Matthew Hirschey, Ph.D."
date: "April 8-9, 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

# Welcome back

# "The simple graph has brought more information to the data analyst’s mind than any other device."  -- John Tukey

## Visualize Data with
```{r ggplot_logo, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", out.height = '100%'}
url <- "https://github.com/tidyverse/ggplot2/raw/master/man/figures/logo.png"
knitr::include_graphics(url)
```
<font size = "1">[https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org)</font>

## A basic plot using `ggplot` in R
We ended the last class with:

```{r ggplot_basic, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
ggplot(heart, aes(x = sex, y = chol)) +
  geom_point(color = "darkblue", size = 3) +
  labs(x = "sex", 
       y = "Cholesterol", 
       title = "Cholesterol values from the heart disease dataset", 
       caption = "Data from Kaggle | Plot from @matthewhirschey") +
  theme_minimal() +
  NULL
```

## which built this...
```{r ggplot_basic2, message=FALSE, warning=FALSE, echo=FALSE}
#need to parameterize this; embed rmd here?; but still need to parameterize that doc.
ggplot(heart, aes(x = sex, y = chol)) +
  geom_point(color = "darkblue", size = 3) +
  labs(x = "sex", y = "Cholesterol", title = "Cholesterol values from the heart disease dataset", caption = "Data from Kaggle | Plot from @matthewhirschey") +
  theme_minimal() +
  NULL
```

# Basic ggplot2
Any questions from last week?

## Basics of a ggplot code {.build}

Below is an example of the most basic form of the ggplot code  
  
`ggplot(data = dataframe) +`   
  `geom(mapping = aes(x, y))`  
  
Take a moment to look back at the code template.  You can see that in that code we assigned a `dataset` and the information we needed to `map` it to a type of plot

## Basics of a ggplot code {.build}
  
`ggplot(data = `r dataframe_name`) +`  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  
  
```{r simple_plot, echo = FALSE, warning = FALSE}
df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 
```

## Build a `ggplot` {.build}
- "Initialize" a plot with `ggplot()`  
  
`ggplot(data = `r dataframe_name`) +`  
  
- Add layers with `geom_functions`  
  
`geom_point(mapping=aes(x = `r df_numeric1_name`, y = `r df_numeric2_name`))`  

## Notes on `ggplot` style {.build}
- Generally, different people have **strong** opinions about style and data visualization  
  
- Data visualization is a rich and complex area of study and is beyond the scope of this introductory course  
  
**That being said, here are a few style tips:**    
- While you can put the `+` at the beginning of the next line, it is generally put at the end of the previous line    
  
- Arguments do not need to be explicit: this chunk of code is equivalent to the previous chunk (w/ data, mappings)  
`ggplot(`r dataframe_name`) +`    
`geom_point(aes(x = `r df_numeric2_name`, y = `r df_numeric2_name`))`  

## How are these two plots similar?
```{r double_plot, echo = FALSE, message=FALSE, warning=FALSE}
plot1 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 

plot2 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_smooth() 

plot1 + plot2
```

## The `geom` is different between these plots
`geom` is short for geometric object, which is the visual object used to represent the data  
    
`plot1 <- ggplot(`r dataframe_name`) +`  
  `geom_point(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   
  
`plot2 <-  ggplot(`r df_numeric1_name`, `r df_numeric2_name`) +`  
  `geom_smooth(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   

## Choosing your (geom)itries
Different data types require different plot types.  
  
When plotting your data, it is often helpful to take a glimpse at the data you intend to plot to know what kinds of variables you will be working with

`glimpse(`r dataframe_name`)`
```{r glimpse_again, exercise=TRUE}

```

## Geoms 
So now that you know your variable types, how do you know what geoms to use??  
  
Use the following resources to match your data type to the appropriate `geoms`

----
<a href="https://rstudio.com/resources/cheatsheets" target="_blank">https://rstudio.com/resources/cheatsheets/</a><br>
<center><iframe src="https://rstudio.com/resources/cheatsheets"></iframe></center>

----
<a href="https://www.data-to-viz.com/" target="_blank">https://www.data-to-viz.com/</a><br>
<center><iframe src="https://www.data-to-viz.com/"></iframe></center>

## Change this scatterplot code to draw boxplots {.smaller} 
`ggplot(`r dataframe_name`) + geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  
```{r boxplot_activity, exercise = TRUE}

```
  
Use the cheatsheet. Try your best guess.  

## What will this code do?
  
`ggplot(`r dataframe_name`) +`   
`geom_boxplot(aes(x=`r df_char1_name`,y=`r df_numeric1_name`)) +`  
`geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  

## Try including two `geoms` for yourself
```{r 2geom_activity, exercise = TRUE}

```
  
**Each new geom adds a new layer**  

## Aesthetics
Everything up to this point gets you a basic graph; but what about colors, shapes and overall style?  
  
You can change 5 basic aesthetics  
1. **Color**- changes the outline color of your datapoints  
2. **Size** - choose the size of the datapoint  
3. **Shape** - choose a pre-defined shape  
4. **Alpha**- changes the transparency of each point  
5. **Fill**- changes the fill color of your points  
  
Go to code/  
Open 04_ggplot2.Rmd  
Complete the exercise.    
  
## Mapping these aesthetics to data
Beyond simply changing the size or color of the variables in your plot, you can encode more information by mapping these values to data in your data set.  
  
Go to code/  
Open 05_aes.Rmd  
Complete the exercise.  
  
## Global vs Local {.build}
In ggplot2, we have the options to set mappings globally or locally. Setting a mapping globally means to set those values in the original ggplot function.  
  
Example: Earlier in class you made this graph:

`ggplot(`r dataframe_name`) +`  
  `geom_jitter(aes(x=`r df_char1_name`, y=`r df_numeric1_name`))+`  
  `geom_boxplot(aes(x=`r df_char1_name`, y=`r df_numeric1_name`))`  
  
## Global vs Local
```{r global_v_local_concepts, echo = FALSE, warning = FALSE}
ggplot(df_input) +
  geom_jitter(aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_boxplot(aes(!!sym(df_char1_name), !!sym(df_numeric1_name)))
```

## Global vs Local
However, if we map our x and y values in the ggplot function we find that we generate the same graph  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot()`  
```{r global_v_local, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot()
```

## Global vs Local {.build}
This is because when you set the aes mappings in the original `ggplot` function you are setting the `aes` **globally**.  
This means all the functions afterwards will inherit that mapping. So in our example, this means that both the jitter and boxplot geoms know to graph the same information  
  
You can also set aes values **locally** within the geom function. Doing so will only change the values in that geom

## Global vs Local
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`))`      
```{r global_v_local1, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name)))
```

## Global vs Local {.build}
```{r global_v_local2, echo=FALSE}
mean <- mean(df_numeric2_vec)
sd <- sd(df_numeric2_vec)
```

Data can also be set locally or globally. For this example, let's filter our original data first using the `dplyr::filter` function
  
`df_filter <- `r dataframe_name` %>% filter(`r df_numeric2_name` > `r round(mean + 2*sd)`)`  
  
*this number is two standard deviations above the `mean` value of `r df_numeric2_name`

## Global vs Local {.build}
Now, let's identify only the `r dataframe_about` in our data that are outliers, more than 2SD above the mean, by setting data locally in a new geom  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`   
`geom_label(data=df_filter, aes(label=`r df_id_name`))`  

## Global vs Local {.build}
```{r global_v_local3, echo = FALSE, warning = FALSE}
df_filter <- df_input %>% filter(!!sym(df_numeric2_name) > round(mean + 2*sd))

ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  geom_label_repel(data = df_filter, aes(label = !!sym(df_id_name)))
```

## Global vs Local {.build}
You notice we have to indicate the new dataset, but because it has the same x and y values, we did not need to set those mappings  
  
Go to code/  
Open 06_global_v_local.Rmd  
Complete the exercise to practice mapping locally and globally.   

## Labels and Legends
Several options exist to change the default labels and legends. Recall, this code:  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`))`   

## Labels and Legends {.build}
```{r labs, echo = FALSE, warning = FALSE}
original_plot <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name)))
original_plot
```
  
But it has two problems:  
1. The x-axis label is redundant  
2. The figure legend is also redundant  

## Change labels using `labs`
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(x ="") #blank quotes removes the label`

## Change labels using `labs`
Gave us this plot:  
```{r labs_fix, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(x = "")
```

## Change legend using `guides`
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(x ="") #blank quotes removes the label +`  
`guides(color = "none")`

## Change legend using `guides` 
```{r guides, echo = FALSE, warning = FALSE}
lab_plot <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(x = "") + 
  guides(color = "none")
original_plot + lab_plot
```

## Faceting {.build}
Faceting allows you to create multiple graphs side by side in one panel. Especially useful when you want to see the data together, but not on top of each other  
  
For example:  
  
`ggplot(`r dataframe_name`) +`  
`geom_point(aes(x=`r df_char1_name`, y=`r df_numeric1_name`)) +`  
`facet_grid(cols = vars(`r df_char2_name`))`

## Faceting
```{r facetting, echo = FALSE, warning = FALSE}
ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_point() +
  facet_grid(cols = vars(!!sym(df_char2_name)))
```
  
*This is especially useful for exploratory data analysis

## Themes {.build}
You can change almost everything you see on your chart, but a lot of the things you may look to change are part of the "theme"  
  
Here we are going to change some features about our title text:  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12))` 

## Themes {.build}
```{r themes, echo = FALSE, warning = FALSE}
theme1 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12)) 
original_plot + theme1
```

## Themes {.build}
Next, let's change the aesthetics of our legend box  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"))`   

## Themes {.build}
```{r themes2, echo = FALSE, warning = FALSE}
theme2 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black")
  ) 
theme1 + theme2
```

## Themes {.build}
Finally, let's change the legend position  
  
`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme(plot.title = element_text(face = "bold", size = 12),`  
`legend.background = element_rect(fill="gray", colour="black"),`  
`legend.position = "bottom"))` 

## Themes {.build}
```{r themes3, echo = FALSE, warning = FALSE}
theme3 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme(plot.title = element_text(face = "bold", size = 12), 
        legend.background = element_rect(fill="gray", colour="black"),
        legend.position = "bottom"
  ) 
theme2 + theme3
```

## Themes {.build}
Pre-set themes also exist as an easy way to change the entire theme of your graph quickly. They can also be combined with custom theme settings  

`ggplot(`r dataframe_name`, aes(x=`r df_char1_name`, y=`r df_numeric1_name`) +`  
`geom_jitter() +`    
`geom_boxplot(aes(color = `r df_char1_name`)) +`  
`labs(title = "My first plot") +`  
`theme_minimal()`  

## Themes {.build}
```{r themes4, echo = FALSE, warning = FALSE}
theme4 <- ggplot(df_input, aes(!!sym(df_char1_name), !!sym(df_numeric1_name))) +
  geom_jitter() +
  geom_boxplot(aes(color = !!sym(df_char1_name))) +
  labs(title = "My first plot") +
  theme_minimal()
theme1 + theme4
```

## Saving plots using `ggsave` {.build}

If you make a plot there are a few ways to save it, though the simplest is to use `ggsave`  

`ggsave("ggsaveexample.png", plot = last_plot())`  

You can change the type of file you save or the size.  
  
`ggsave("ggsaveexample.pdf", plot = my_awesome_object, width = 6, height = 6, units = "cm")`  

## Saving plots using ggsave
Where does it save??  
  
`getwd()`

## Putting it all together
Go to code/  
Open 07_ggplot_together.Rmd  
Complete the exercise to put all these ggplot skills to work.  

# R <br> Importing
Any questions?


# R <br>Strings
Any questions?

## Strings hold text
A string is what we store text within. It can be either:  
- A single word: `"awesome"`  
- A sentence: `"this class is awesome"`  
- A combination: `c("blue", "is my favorite", "color")`  
  
Any of these can be stored as object, which we call strings. 

```{r string_exercise, exercise=TRUE}

```

## Dealing with strings {.build}
Dealing with character strings is a bit different than dealing with numbers in R?   
  
* Sort?  
* Select?  
* Change?  
  
Fortunately, the `tidyverse` as a package called `stringr` for dealing with them. 

## Stringr
<center><iframe src = "https://stringr.tidyverse.org"></iframe></center>

## Counting string {.build}
```{r setup_count, include=FALSE}
df_input2 %>% count(health_status, sort = TRUE) %>% slice(2) %>% select(1) %>% pull()
#restart here
```

`str_count()` is a function we can use to count the number of rows that match a particular pattern.  
  
The output with either be 1 (match), or 0 (no match)  

<div class= "red2"> Example </div>  
In this code:  
- string we want to evaluate is ``r dataframe_joined_name`$`r df_joined_string3_name``  
- pattern we want to count `"High Cholesterol"`  

`str_count(heart_joined$health_status, "High Cholesterol")`

## Counting
`str_count(heart_joined$health_status, "High Cholesterol")`

```{r str_count example, echo= FALSE, warning = FALSE}
str_count(heart_joined$health_status, "High Cholesterol")
```

## Summarizing our counts
A bunch of 0 and 1 are not incredibly useful.

But since R is good at adding, we can simply wrap the previous expression in `sum()`  

Try it below:  
```{r sum_string, exercise = TRUE}

```

## Matching subsets of strings {.build}

We previously matched the entire string "High Cholesterol"  
But we can use the same function to detect patterns within longer strings.  
  
Let's look for how many patients take a statin of any kind using
`str_count(heart_joined$medication_hx, "statin")` 

## What about subsets of strings?  
`str_count(heart_joined$medication_hx, "statin")` 

What does the ouput mean?  
```{r str_count_exercise, exercise = TRUE}

```

## A note about string patterns

When using a stringr function, you may get an output saying a string pattern doesn't exist. If you know for sure it does, <div class="blue"> double check capitalization </div>.  
  
**The string must match exactly, or it will not be found!**

## stringr Exercise
How many people having an `"auntie" or "aunt"` in their health history?  
  
Go to code/  
Open 09_stringr.Rmd  
Complete the exercise. 

## Using Regular Expressions
That solution worked in this case, but was not very elegant, and might not work for all cases (what if there was a 'great aunt' in the list?)  

Or here is a more specific case for this data set.  
  
How many patients have a father with a history of disease? But we don't want to include grandfathers in the results. 
  
We can use something called **Regular Expressions**, aka **Regex**, to solve this

## Using Regular Expressions {.build}
Think of regex as a separate language, with it's own code, syntax, and rules.  
  
Regex rules allow complex matching patterns for strings, to ensure matching *exactly* the content desired
  
It is far too complex to cover in its entirely here, but here is one specific example.  
  
**GOAL**: identify all of the patients that have a father with a history of disease, but excluding grandfathers in the results. 
  
## Regular Expression Example {.build}
<div class="red2"> Example </div>  
We want to start with recognizing `father`.  

But then we want to make sure that we capture both `Father` and `father`. To accept either case f in the first spot we add `(F|f)`, so now our regex looks like `(F|f)ather`  
  
Lastly, we want this pattern to appear at the beginning of the word, so we add the regex `^` symbol.  

Our completed regex looks like:  
  
`str_count(heart_joined$family_history, "^(F|f)ather")` 

```{r regex1, exercise=TRUE}


```
## Regex Exercise
Go to code/  
Open 09_stringr.Rmd  
Complete the exercise to count mothers.    

## Regex resources
- The stringr cheatsheet shown above is a great resource  
- Here the tidyverse website with a link to its vignette  
  
<center><iframe src="https://stringr.tidyverse.org/articles/regular-expressions.html"></iframe></center>  

## Detecting strings {.build}
In addition to counting, we can use another function `str_detect()` to logically evaluate a character string.  
  
Because this logically evaluates an expression, the output is either `TRUE or FALSE`  
  
Practially, `str_detect` is used to detect the presence or absence of a pattern in a string  

## Logic Evaluation 
<div class= "red2"> Example </div>  
Find the patients with diabetes using the following code  
  
`str_detect(heart_joined$health_status, "Diabetic")`  
  
```{r strdetect, exercise=TRUE}

```

## Modifiying strings with `str_replace()` {.build}
In the health_status column we have:  
-`"Diabetic"`  
-`"High Cholesterol"`  
-`"Normal blood sugar and cholesterol"`  
  
But let's say we want to simplify healthy individuals to `"Normal"`

`str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "Normal")`

```{r str_replace, exercise=TRUE}

```

## Modifiying strings with `str_replace()` {.build}
We use this same code to  modify the `health_status` column by assigning it to the same variable     
  
`heart_joined$health_status <-`  
       `str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "normal")`

```{r str_replace in col, warning = FALSE, echo=FALSE}
heart_joined$health_status <- 
  str_replace(heart_joined$health_status, "Normal blood sugar and cholesterol", "normal")

head(heart_joined$health_status, n = 10)
```

## Using `stringr` with `dplyr` {.build}
We can use `stringr` functions in tandem with `dplyr` functions.  
  
<div class= "red2"> Example </div>  
We want to make a logical variable (`TRUE`/`FALSE`) that tells us if a patient has a normal health history using  

`heart_joined2 <- heart_joined %>% mutate(healthy = str_detect(health_status, "normal"))` 

```{r stringr with dplyr, echo=FALSE, warning=FALSE}
heart_joined2 <- mutate(heart_joined, healthy = str_detect(health_status, "normal"))

head(heart_joined2$healthy, n = 10)
```

## Regex Exercise
Go to code/  
Open 09_stringr.Rmd  
Complete the exercise to count mothers.   

# R <br> Markdown
Any questions?

## Scientific Reproducibly and Provenance with
```{r rmarkdown_logo, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/rstudio/hex-stickers/raw/master/PNG/rmarkdown.png"
knitr::include_graphics(url)
```
<font size = "1">[www.rstudio.com](www.rstudio.com)</font>

## R Markdown
Plain text file with 3 types of content  
```{r rmarkdown_3parts, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown.png"
knitr::include_graphics(url)
```

## Use the 'knit' button to render a report
```{r rmarkdown_knit_button, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/96ec0c54c6d64ea2ec3665db9b3b781962ff6339/5cee1/lesson-images/how-3-output.png"
knitr::include_graphics(url)
```

## How it works
```{r rmarkdown_knit, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/61d189fd9cdf955058415d3e1b28dd60e1bd7c9b/b739c/lesson-images/rmarkdownflow.png"
knitr::include_graphics(url)
```
  
When you run render, R Markdown feeds the .Rmd file to knitr ⧉, which executes all of the code chunks and creates a new markdown (.md) document which includes the code and it’s output.  
  
The markdown file generated by `knitr` is then processed by `pandoc` which is responsible for creating the finished format.  
  
This may sound complicated, but R Markdown makes it extremely simple by encapsulating all of the above processing into a single render function.  
  
## Key `knitr` points {.build}
- `Knitr` runs the document in a fresh R session, which means you need to load the libraries that the document uses **in the document**  
  
- Objects made in one code chunk will be available to code in later code chunks, but not before  
  
- For example, first create `r dataframe_name` and then using `dplyr::left_join` you create `r dataframe_joined_name`, `r dataframe_name` will be available later on in the document to do this. **However, you cannot use `r dataframe_joined_name` in a code chunk before you make it, even if it is available in your environmnet**  
  
- To keep this straight, just think (and code) in sequential chunks

## Rmarkdown -- The definitive Guide
[https://bookdown.org/yihui/rmarkdown/](https://bookdown.org/yihui/rmarkdown/)  
<center><iframe src="https://bookdown.org/yihui/rmarkdown/"></iframe></center>

# Rmarkdown <br>Text
Any questions?  

## Markdown is a simplifed language to format text
```{r rmarkdown_text, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://d33wubrfki0l68.cloudfront.net/59f29676ef5e4d74685e14f801bbc10c2dbd3cef/c0688/lesson-images/markdown-1-markup.png"
knitr::include_graphics(url)
```

## Markdown can 'markup' text to be {.build}
- **bold**  
- _italics_  
- `code`  
- unformatted text  
- ### Medium headers (3)  
- ###### Small headers (6)  
- Bullets (like this list)  
- [Links, such as to the github repository storing class material](https://github.com/matthewhirschey/)  
- ...and many more
  
## Rmarkdown cheatsheet & quick reference
<center><iframe src="https://rmarkdown.rstudio.com/lesson-15.html"></iframe></center>
  
# Rmarkdown <br>Code
Any questions?  

## Rmarkdown Code
```{r rmarkdown_code, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Shortcut
```{r rmarkdown_code_shortcut, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code-shortcut.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk1, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk1.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk2, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk2.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk3, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk3.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk4, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk4.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk5, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk5.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk6, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk6.png"
knitr::include_graphics(url)
```

## Rmarkdown Code Chunk Options
```{r rmarkdown_chunk7, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk7.png"
knitr::include_graphics(url)
```

## Clean-up the Rmarkdown exercise
1. Go to code/  
2. Open 10_rmarkdown.Rmd  
3. Inspect the code starting with the set-up chunk. Below it, you’ll see code that imports data, stores objects, and creates a graph.  
4. Knit the document and see the results; pretty messy, huh?  
5. Use what you've learn to execute the code, but not include the results of the set-up chunk and data import chunk.  
6. Replace every **BOLD** with inline R code  
7. Remove the {calculations} chunk so it is not included with the output  
8. Change the {graph} chunk so that only the output of the plot is shown, but not the code.
9. Re-Knit the document  
10. **Beautiful!**  

# Rmarkdown <br>YAML
Any questions?  

## Rmarkdown header is called 'YAML'
```{r rmarkdown_yaml, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-yaml.png"
knitr::include_graphics(url)
```

## Rmarkdown output is defined here
```{r rmarkdown_output, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-output.png"
knitr::include_graphics(url)
```

## Rmarkdown formats
Recall that Rmarkdown documents can be rendered into several different output file types
<center><iframe src="https://rmarkdown.rstudio.com/formats.html"></iframe></center>

## Rmarkdown Parameters
Parameters of a document are defined in the YAML header, and can pre-populate an Rmarkdown document. To see this in action,  
  
1. Open 11_rmarkdown_params.Rmd.  
2. Click the dropdown menu next to Knit and use Knit with Parameters to render the document.  
3. What happens if you type in a different gender or different age?  

--- 

```{r rmarkdown_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-parameters.png"
knitr::include_graphics(url)
```

--- 

```{r rmarkdown_using_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-using-parameters.png"
knitr::include_graphics(url)
```

## Take another look
<h1>Re-inspect 11_rmarkdown_params.Rmd.</h1>
<br>
<br>
Where were the parameters called in the code?

# Concluding Thoughts

## Data science enables {.build}
- Data science enables data-driven information gathering and hypothesis generation  
-- Scientific Research  
-- Reviews  
  
- Data science enables the ability to ask new types of questions  
  
- Process-centric, not necessarily question-centric  
  
- Making things computable makes them actionable at zero marginal cost.  
  
- Workflows save time, achieve reproducibility  

## Cheatsheet Resources
<a href="https://rstudio.com/resources/cheatsheets" target="_blank">Cheatsheets</a><br>
<center><iframe src="https://rstudio.com/resources/cheatsheets"></iframe></center>

## Stackoverflow Resources
<a href="https://stackoverflow.com/" target="_blank">Stackoverflow.com</a><br>
<center><iframe src="https://stackoverflow.com"></iframe></center>

## R4DS Resources
<a href="https://r4ds.had.co.nz" target="_blank">R for Data Science</a><br>
<center><iframe src="https://r4ds.had.co.nz"></iframe></center>

## Tidytuesday Resources
<a href="https://github.com/rfordatascience/tidytuesday" target="_blank">Tidytuesday</a><br>
<center><iframe src="https://github.com/rfordatascience/tidytuesday"></iframe></center>

## Visualization Resources: Tidytuesday
<a href="https://github.com/rfordatascience/tidytuesday" target="_blank">Tidytuesday</a><br>
<center><iframe src="https://github.com/rfordatascience/tidytuesday"></iframe></center>

## Acknowledgements {.build}
**Teaching Assistants**  
- Allie Mills, Ph.D.  
- Akshay Bareja, D.Phil.  
  
**Inspiration, ideas, packages, code**  
- R4DS (Garrett Grolemund and Hadley Wickham)  
- Mine Çetinkaya-Rundel (datasciencebox.org)  
- Chester Ismay and Albert Y. Kim (Modern Dive)  
- Garrett Grolemund  (Remastering the Tidyverse)  
- Tidyverse devs and community  
- Rstudio  

# Thank you
Any questions?