---
title: "Tidybiology +DS: Session 1 "
subtitle: "Introduction to Data Science in Health Care using R"
author: "Matthew Hirschey, Ph.D."
date: "April 8-9, 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(twitterwidget)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Doctors make decisions based on symptoms 
```{r intro, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/701-FJcjLAQ/500x500"
knitr::include_graphics(url)
```
<font size = "1">Photo by [National Cancer Institute on Unsplash](https://unsplash.com/@nci)</font>

## New digital healthcare era introduces new decision-making challenges {.build}
**Volume**  
- Data collection & storage allows access to huge amounts of medical information  
  
**Ubiquity**  
- Data are available anywhere across geography, social, and economic classes   
  
**Latency**  
- Technology facilitates no delay in access to data    
  
## How do you make better health care decisions? {.build}
**Data-driven** decision making!  
  
```{r hierarchy, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>
  
## Emerging field of Data Science
```{r venn1, echo=FALSE, fig.cap="", out.height= '100%'}
url <- "https://images.squarespace-cdn.com/content/v1/5150aec6e4b0e340ec52710a/1364352051365-HZAS3CLBF7ABLE3F5OBY/ke17ZwdGBToddI8pDm48kB2M2-8_3EzuSSXvzQBRsa1Zw-zPPgdn4jUwVcJE1ZvWQUxwkmyExglNqGp0IvTJZUJFbgE-7XRK3dMEBRBhUpxPe_8B-x4gq2tfVez1FwLYYZXud0o-3jV-FAs7tmkMHY-a7GzQZKbHRGZboWC-fOc/Data_Science_VD.png?format=1500w"
knitr::include_graphics(url)
```
<font size = "1">http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram</font>

## Venn Diagram of Data Science v2.0
```{r venn2, echo=FALSE, fig.cap="", out.height= '100%'}
url <- "https://3.bp.blogspot.com/-bvQxcwfqATQ/V-E_uTBc4VI/AAAAAAAAMGQ/Qa1Ntef-rs0E-mWx5pkVu-CPlREdvD0TwCLcB/s1600/VennDiagram2.png"
knitr::include_graphics(url)
```
<font size = "1">Joel Grus via KDnuggets</font>

## Overall goal is Knowledge Generation
```{r knowledge, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://upload.wikimedia.org/wikipedia/commons/0/06/DIKW_Pyramid.svg"
knitr::include_graphics(url)
```
<font size = "1">https://en.wikipedia.org/wiki/DIKW_pyramid</font>

## Several Approaches to Knowledge Generation {.build}
- New tools allow data interrogation more easily than ever before 

```{r approaches, echo=FALSE}
tweet <- twitterwidget('1125268670324695041')
```
<center>`r tweet`</center>

## World's most popular programming languages
```{r popular, echo = FALSE}
excel <- tibble(
  name = c("Excel", "Java", "C", "C++", "Python"), 
  num = c(100000000, 9000000, 6000000, 4000000, 3000000)
)
excel_plot <- ggplot(excel) +
  geom_col(aes(x = fct_rev(fct_reorder(name, num)), y = num), fill = "navy") +
  labs(x = "", y = "Number of Users (Million)") +
  scale_y_continuous(labels = c("0", "25", "50", "75", "100")) +
  theme_minimal()
excel_plot
#need to add source
```

# R <br> language
Any questions?

## R is a language
```{r language, echo=FALSE, fig.cap=""}
url <- "https://source.unsplash.com/ZzWsHbu2y80/256x455"
knitr::include_graphics(url)
```
<font size = "1">Photo by Hannah Wright on Unsplash</font>

## R has values
- 1
- "North Carolina"
- "2020-04-08"

## R has objects
- A name without quotes  
- Assigned using `<-` (looks like an arrow pointing left)  
- Can be a value, object, or function result  

## Try assigning an object 
1. Assign an object; **remember, no quotes on name**  
`name <- 4`  
  
2. Return that object by typing its name  
`name`  
  
Try this in the code chunk below, then hit "Run Code"  

```{r my_first_object, exercise=TRUE}

```

## R has functions
- A name without quotes  
- Followed by () to run the function  
- Optional arguements: values, objects, or function results  
- `round(x, digits = 3)`

Try this in the code chunk below, then hit "Run Code"  

```{r my_first_fun, exercise=TRUE}
round(pi, digits = 3)

```

## Which of these are numbers?

```{r num_quiz, echo=FALSE}
question("Which of these are numbers?",
  answer("1", correct = TRUE),
  answer('"1"', message = "Because it has quotes, it is an string"),
  answer('"one"', message = "Because it has quotes, it is an string"),
  answer("one",  message = "R recognizes this as an object"), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```

## Which of these will work?
Suppose `one <- 1`
```{r one_quiz, echo=FALSE}
question("Which of these will work?",
  answer("log(1)", correct = TRUE),
  answer('log("1")', message = "You cannot log transform a word (string)"),
  answer('log("one")', message = "You cannot log transform a word (string)"),
  answer("log(one)", correct = TRUE), 
  allow_retry = TRUE, 
  random_answer_order = TRUE
)
```
Try it for yourself!  
  
```{r scratchpad, exercise=TRUE}
one <- 1
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r kable}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  knitr::kable()
```

## Data are stored in tables and dataframes
Data stored in a dataframe are conceptually equivalent to a spreadsheet with rows and columns  
  
This is a sample from the ``r dataframe_name`` dataset  
  
```{r tibble}
df_input %>% 
  select(1:5) %>% 
  slice(1:5) %>% 
  print(as_tibble())
```

## Extract or create new objects
```{r extract_setup, include=FALSE}
col <- sample(colnames(df_input), 1)
```

You can call a single part of the data frame  
  
`r dataframe_name`$`r col`

```{r extract, echo = FALSE, warning = FALSE}
df_input %>% 
  select(!!col)
```

## Extract or create new objects 
  
Write the R code required to extract a variable from the `r dataframe_name` dataset:  
  
Remember, the format is: `r dataframe_name`$`r col`  
  
```{r extract_exercise, exercise=TRUE}

```

## Extract or create new objects 
  
You can also save a part of the dataframe as an object for later use  
  
``r col` <- `r dataframe_name`$`r col``  
  
In the code chunk below:  
1. On the first line, write the R code to save a single column to a new object   
2. On the second line, type the object name - this will print out the new object  
3. Run the code  
  
```{r create_exercise, exercise=TRUE}

```

# R <br>Integrated Development Environment
Any questions?

## R
Statistical Programming Language
```{r r_exercise, exercise=TRUE}

```
  
```{r r_image, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/Vb8pR6cKCRI/300×200"
knitr::include_graphics(url)
```
<font size = "1">Photo by [Photo by Evan Hein on Unsplash](https://unsplash.com/@evanheinphoto)</font>

## Rstudio
Integrated Development Environment

```{r r_de, echo=FALSE, fig.cap="", out.height = '100%'}
url <- "https://source.unsplash.com/XuDPnpox8tc/300×200"
knitr::include_graphics(url)
```
<font size = "1">[Photo by Jaromír Kavan on Unsplash](https://unsplash.com/@jerrykavan)</font>

---
[Open Rstudio](http://www.rstudio.cloud)  
<iframe src="http://www.rstudio.cloud"></iframe>

## Rstudio Demonstration
Go to code/  
Open 01_r_demo.Rmd  
Follow along.  

# R <br>Markdown
Any questions?

## R Markdown Introduction
Go to code/  
Open 02_rmd_exercise.Rmd  
Read through the file and do everything it tells you to do.  

## R Markdown
An authoring format for Data Science.  
```{r markdown, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/markdown.png"
knitr::include_graphics(url)
```

# R <br>Packages
Any questions?

## R Packages
The R language contains thousands of functions, data sets, and help pages.  
- but only a few hundred are included when you download R  
  
**This is called 'Base R'**  
  
The other functions, data sets, and help pages are grouped into collections known as packages that you can choose to download or not download.  

## “Verbs” (i.e. functions) act on data

`do_this(to_that)`  
  
`do_this(to_that, using_these)`  
   
*We talked about functions before (e.g. `round(pi, 3)`)*  
  
  
  
  
**Functions are the power of using R**

## Packages contain functions, documentation, data
  
<center><iframe src="https://dplyr.tidyverse.org"></iframe></center>

## CRAN
- Most R packages are stored on CRAN, alongside R.  
- Think of them as optional extensions of the R language.  
  
```{r cran, echo=FALSE, fig.cap="", out.height = '66%', out.width= '66%'}
url <- "https://gist.githubusercontent.com/daroczig/3cf06d6db4be2bbe3368/raw/b66b0531fb1b86d3e04a003b2e105ad4f147900e/number-of-submitted-packages-to-CRAN.png"
knitr::include_graphics(url)
```
<font size = "1">Image by [daroczig](https://gist.github.com/daroczig/3cf06d6db4be2bbe3368)</font>
  
## Using Packages (Part I)
  
1. `install.packages("tidyverse")`  
  
Do this 1 time per computer.  
  
This command will install the package into your instance of R, whether it is local, on a server, or in the cloud. This is required to use the functions in a package.  
  
## Tidyverse
  
<center><iframe src="https://www.tidyverse.org"></iframe></center>

## Tidyverse Pop Quiz!

The tidyverse contains the following packages (ggplot2, dplyr, tidyr, readr, purrr, tibble, hms, stringr, lubridate, forcats, DBI, haven, httr, jsonlite, readxl, rvest, xml2, modelr, tidyverse).   
  
  
  
  
**How would you  install them?**

## Tidyverse Pop Quiz! {.build}
`install.packages("ggplot2")`
`install.packages("dplyr")`
`install.packages("tidyr")`
`install.packages("readr")`
`install.packages("purrr")`
`install.packages("tibble")`
`install.packages("hms")`
`install.packages("stringr")`
`install.packages("lubridate")`
`install.packages("forcats")`
`install.packages("DBI")`
`install.packages("haven")`
`install.packages("httr")`
`install.packages("jsonlite")`
`install.packages("readxl")`
`install.packages("rvest")`
`install.packages("xml2")`
`install.packages("modelr")`
`install.packages("broom")`  
  
  
  
Better:  
`install.packages("tidyverse")`  
An R package that serves as a short cut for installing and loading the components of the tidyverse.

## Using Packages (Part II)
  
1.`install.packages("tidyverse")`  
  
Do this 1 time per *computer*.  
  
  
2.`library(tidyverse)`  
  
Do this 1 time per *session*  
  
## Using Packages (Part II) {.build}
**Downloading a package isn't the same as using it.**  
  
If you'd like to use an R package, you need to tell R.  
You do that by running the command `library`, again followed by parentheses and the package name.  
  
`library(package_name)`  
  
This command loads all of the  functions, data sets, and help pages that come with the package into your R session, where you can use them.  
  
*If you close R, you'll need to reload the package with library() if you want to use it again.*  
  
  
## Using Packages (Part III)
The setup chunk is always run once before anything else  

```{r setup_chunk, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/setup.png"
knitr::include_graphics(url)
```

## Set-up Chunk Exercise
Add a setup chunk (as shown below) to the top of `02_rmd_exercise.Rmd`.  
Use it to load the tidyverse package (*remember to run this chunk*)  
Then uncomment and run the final code chunk at the bottom of your file.  
  
```{r setup_chunk2, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/setup2.png"
knitr::include_graphics(url)
```

## Tidyverse 
Tidyverse is one suite of tools for data science  
  
<center><iframe src="https://www.tidyverse.org"></iframe></center>

## Exploratory Data Analysis
```{r r4ds, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/hadley/r4ds/raw/master/diagrams/data-science.png"
knitr::include_graphics(url)
```

## Tidyverse Basic Principles
**IMPORT** (readr):  
- `read_csv()`  
- `read_delim()`  
  
**TIDY & TRANSFORM** (dplyr):  
- `mutate()` adds new variables that are functions of existing variables  
- `select()` picks variables based on their names.  
- `filter()` picks cases based on their values.  
- `summarise()` reduces multiple values down to a single summary.  
- `arrange()` changes the ordering of the rows.  
  
**VISUALISE** (ggplot): creating graphics, based on ‘The Grammar of Graphics’  
- `aes()`  
- `geom_x()` + layers   
  
**MODEL** (broom):   
- `tidy()`, `glance()`, `augment()`  

----
```{r pipe, echo=FALSE, fig.cap="", out.width='77%'}
url <- "https://s3-us-west-2.amazonaws.com/collections.lacma.org-images/remote_images/ma-150089-WEB.jpg"
knitr::include_graphics(url)
```

## `magrittr` package
```{r magrittr, echo=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/tidyverse/magrittr/raw/master/man/figures/logo.png"
knitr::include_graphics(url)
```

## `magrittr` package
`magrittr` package by Stefan Milton Bache developed the concept of the pipe, which is used heavily in the `tidyverse`  
  
  
```{r magrittr2, echo=FALSE, fig.cap="", out.width='40%'}
url <- "https://github.com/tidyverse/magrittr/raw/master/vignettes/magrittr.jpg"
knitr::include_graphics(url)
```
<br>  
<center><font size = "10">"and then"</font></center>

## The Pipe {.build}
  The “pipe” is a sequence of functions, that are sequentially applied to an object  
```{r wakeup, echo=TRUE, eval=FALSE}
wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  go(to = "work")
```
  
  Alternative nested code:  
```{r wakeup_nested, echo=TRUE, eval=FALSE}
go(eat(put_on(wakeup(self), "clothes"), "breakfast"), to = "work")
```

## The Pipe (Quiz I)
What does this code do? 
  
```{r wakeup_quiz, echo=TRUE, eval=FALSE}
wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  fmk() %>% 
  go(to = "work")
```
  
## The Pipe (Quiz II)
What does this code do? 
  
```{r wakeup_quiz2, echo=TRUE, eval=FALSE}
morning_routine <- wakeup(self) %>%
  put_on("clothes") %>%
  eat("breakfast") %>%
  fmk() %>% 
  go(to = "work")
```

## The Pipe (Quiz III)
What does this code do? 

``r dataframe_name` %>%`  
`select(`r colnames(df_input)[2]`, `r colnames(df_input)[3]`) %>%`  
`group_by(`r colnames(df_input)[3]`) %>%`  
`summarize(mean(`r colnames(df_input)[2]`))`  
  
Try it out here!
```{r pipe_ex, exercise=TRUE}

```

## Writing code IS NOT like drawing an owl
```{r owl, echo=FALSE, fig.cap="", out.width='50%'}
url <- "https://66.media.tumblr.com/tumblr_l7iwzq98rU1qa1c9eo1_500.jpg"
knitr::include_graphics(url)
```

## Writing code IS a step-wise process
```{r lego, echo=FALSE, fig.cap="", out.width='70%'}
url <- "https://github.com/rstudio-education/datascience-box/raw/master/slides/u1_d01-meet-the-toolkit/img/lego-steps.png"
knitr::include_graphics(url)
```
<font size = "1">https://datasciencebox.org</font>

# R <br>Getting to know your data
Any questions?

## ``r dataframe_name`` dataset
**`r dataframe_title`**  
  
- For this class, we will use a dataset called ``r dataframe_name``.  
  
- `r dataframe_source`  
  
- `r dataframe_desc`  
  
## Inspecting your dataframe: dimensions

Use the `dim()` function to see how many rows (observations) and columns (variables) are in ``r dataframe_name``

`dim(`r dataframe_name`)` 
  
Enter your function here:  
```{r dim_exercise, exercise=TRUE}

```

## Inspecting your dataframe: glimpse

Use the `glimpse()` function to see what kinds of variables the ``r dataframe_name`` dataset contains   
  
`glimpse(`r dataframe_name`)`  
  
Enter your function here:  
```{r glimpse_exercise, exercise=TRUE}

```
  
## Basic Data Types in R {.build}

Recall that R has different data types -

**character** - `"a"`, `"tidyverse"`  

**numeric** - `2`, `11.5`  

**integer** - `2L` (the `L` tells R to store this as an integer)  

**logical** - `TRUE`, `FALSE`  

**complex** - `1+4i`  

(**raw**)

You will also come across the **double** datatype. It is the same as **numeric**

**factor**. A **factor** is a collection of *ordered* character variables

## Basic Data Types in R

In addition to the `glimpse()` function, you can use the `class()` function to determine the data type of a specific column

`class(`r dataframe_name`$`r colnames(df_character_select)`)`

```{r class, echo = FALSE, warning = FALSE}
class(df_character_select[[1]])
```
  
Try getting the `class` of a variable:  
```{r class_exercise, exercise=TRUE}

```

## (Re)Introducing `%>%` {.build}

The `%>%` operator is a way of "chaining" together strings of commands that make reading your code easy.  
The following code chunk illustrates how `%>%` works:  
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  

```{r %>%, echo = FALSE, warning = FALSE}
df_input %>%
  select(colnames(df_character_select), colnames(df_numeric_select)) %>% 
  filter(!!rlang::sym(colnames(df_character_select)) == as.character(df_character_filter)) %>% 
  head()
```

## (Re)Introducing `%>%` {.build}

The previous code chunk does the following - it takes your dataset and then "pipes" it into `select()`, and then applies a `filter()` to the data.
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  
  
*the `head` function lists only the top n results -- convenient for long variables*
  
  
**When you see `%>%`, think "and then"**

## (Re)Introducing `%>%` {.build}

The alternative to using `%>%` is running the following code

filter(select(`r dataframe_name`, `r colnames(df_character_select)`, `r colnames(df_numeric_select)`), `r colnames(df_character_select)` == "`r df_character_filter`")

Although this is only one line as opposed to three, it's both more difficult to write and more difficult to read

# R <br>Manipulating your data
Any questions?

## Introducing `dplyr` {.build}
  
dplyr is a package that contains a suite of functions that allow you to easily manipulate a dataset  
  
Some of the things you can do are -  
  
- select rows and columns that match specific criteria  
  
- create new variables (columns)  
  
- obtain summary statistics on individual groups within your datsets  
  
The main verbs we will cover are `select()`, `filter()`, `arrange()`, `mutate()`, and `summarise()`. These all combine naturally with `group_by()` which allows you to perform any operation "by group"

## `select()` specific columns from your dataset {.build}
The most basic `select()` is one where you comma separate a list of columns you want included

For example, if you only want to select the `r colnames(df_character_select)` and `r colnames(df_numeric_select)` columns, run the following code chunk

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`  
    `head()`
    
```{r select_exercise, exercise=TRUE}

```

## `select()` {.build}

If you want to select all columns *except* `r colnames(df_character_select)`, run the following

``r dataframe_name` %>%`  
    `select(-`r colnames(df_character_select)`) %>%`  
    `head(5)`

```{r selectnot_exercise, exercise=TRUE}

```

## `select()` {.build}

Finally, you can provide a range of columns to return two columns and everything in between. For example

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`:`r colnames(df_numeric_select)`) %>%`  
    `head(5)`
    
```{r selectrange_exercise, exercise=TRUE}

```

## `filter()` rows based on certain condition(s) {.build}

The `filter()` verb evalutes a logical statement, and if a row meets the condition of this statement (i.e. is true) then it gets chosen (or "filtered").  
  
  
**All other rows are discarded**  

## `filter()` {.build}

Filtering can be performed on **categorical** data

``r dataframe_name` %>%`     
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`     
    `head(3)`  

```{r filter_exercise, exercise=TRUE}

```

Note that `filter()` only applies to rows, and has no effect on columns

## `filter()` {.build}

Filtering can also be performed on **numeric** data

For example, if you wanted to choose `r colnames(df_numeric_select)` with a value greater than `r df_numeric_filter`, you would run the following. 

``r dataframe_name` %>%`     
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`     
    `head(3)`  

```{r filter_exercise2, exercise=TRUE}

```

## `filter()` {.build}

To filter on multiple conditions, you can write a sequence of `filter()` commands

``r dataframe_name` %>%`       
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`  
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`    
    `head(3)`  

```{r filter_exercise3, exercise=TRUE}

```

## `filter()` {.build}

To avoid writing multiple `filter()` commands, multiple logical statements can be put inside a single `filter()` command, separated by commas

``r dataframe_name` %>%`           
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`",`  
           ` `r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`      
    `head(3)`  

```{r filter_exercise4, exercise=TRUE}

```

## `arrange()` sorts rows {.build}

The input for arrange is one or many columns, and `arrange()` sorts the rows in ascending order i.e. from smallest to largest

For example, to sort rows from smallest to largest `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`   
  `arrange(`r colnames(df_numeric_select)`) %>%`   
  `head(3)`  

```{r arrange_exercise, exercise=TRUE}

```

## `arrange()` {.build}

To reverse this order, use the `desc()` function within `arrange()`

``r dataframe_name` %>%`  
  `arrange(desc(`r colnames(df_numeric_select)`)) %>%`  
  `head(3)`  

```{r arrange_exercise2, exercise=TRUE}

```


## `mutate()` {.build}

The `mutate()` verb, unlike the ones covered so far, creates new variable(s) i.e. new column(s). For example

``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(1)`  
  
The code chunk above takes all the elements of the column `r colnames(df_numeric_select)`, evaluates the square root of each element, and populates a new column called `new_col` with these results
  
## Try `mutate()` to make a new column
  
``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(3)`  
  
```{r mutate_exercise, exercise=TRUE}

```


## `summarise()` and `summarize()` {.build}

`summarize()` is one of the key functions in `dplyr`. It produces a new dataframe that aggregates that values of a column based on a certain condition.  
  
For example, to calculate the mean `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`      
  `summarise(mean(`r colnames(df_numeric_select)`))` 

```{r summarize_exercise, exercise=TRUE}

```

## `group_by()` {.build}

`group_by()` and `summarize()` are a powerful combination of functions to summarize by groups  
  
``r dataframe_name` %>%`  
  `group_by(`r colnames(df_character_select)`) %>%`  
  `summarise(mean(`r colnames(df_numeric_select)`))`  
  
```{r groupby_exercise, exercise=TRUE}

```
  
*remember this pair!  

## Saving a new dataset as an object

If you'd like to save the output of your wrangling as an object, you will need to use the `<-` operator

``r dataframe_name`_new <- `r dataframe_name` %>%`  
            `group_by(`r colnames(df_character_select)`) %>%`  
            `summarise(mean(`r colnames(df_numeric_select)`))`  
  
*Assigning the object the same name (i.e. ``r dataframe_name``) will overwrite the object  
  
```{r df_object_exercise, exercise=TRUE}

```
  
## Saving a new dataset as a file

To save ``r dataframe_name`_new` as a new file (e.g. csv), run the following:  
  
`write_csv(`r dataframe_name`_new, "`r dataframe_name`_new.csv")`
  
  
# R <br>Data Science Workflow
Any questions?

## Visualizing our dataset
  
Go to code/  
Open 03_rmd_practice.Rmd  
Complete the activity.
  
# R <br>Help
Any questions?

## Resources
- Rstudio package 'cheatsheets'  
- Package function help pages: `?mean`, or navigate to the Help tab and search there  
- Run the following to access the Dplyr vignette: `browseVignettes("dplyr")`  
- Stackoverflow: [https://stackoverflow.com](https://stackoverflow.com)
- R for Data Science, by Grolemund & Wickham [https://r4ds.had.co.nz/index.html](https://r4ds.had.co.nz/index.html)  

## Acknowledgements {.build}
**Teaching Assistants**  
- Allie Mills, Ph.D.  
- Akshay Bareja, D.Phil.  
  
**Inspiration, ideas, packages, code**  
- R4DS (Garrett Grolemund and Hadley Wickham)  
- Mine Çetinkaya-Rundel (datasciencebox.org)  
- Chester Ismay and Albert Y. Kim (Modern Dive)  
- Garrett Grolemund  (Remastering the Tidyverse)  
- Tidyverse devs and community  
- Rstudio  

# Thank you
Any questions?

