---
title: "Introduction to Summarise/Summarize and Group_By"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## `summarise()` and `summarize()` {.build}

`summarize()` is one of the key functions in `dplyr`. It produces a new dataframe that aggregates that values of a column based on a certain condition.  
  
For example, to calculate the mean `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`      
  `summarise(mean(`r colnames(df_numeric_select)`))` 

```{r summarize_exercise, exercise=TRUE}

```

## `group_by()` {.build}

`group_by()` and `summarize()` are a powerful combination of functions to summarize by groups  
  
``r dataframe_name` %>%`  
  `group_by(`r colnames(df_character_select)`) %>%`  
  `summarise(mean(`r colnames(df_numeric_select)`))`  
  
```{r groupby_exercise, exercise=TRUE}

```
  
*remember this pair! 