---
title: "Importing data into R"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Importing data into R {.build}

The `readr` package (found in the `tidyverse` collection) contains a number of useful functions of the form `read_*` to import data. For example, if you have a `.csv` file, you would use the `read_csv` function  
  
The dataset provided to you is a cleaned R-specific document. But you will never find this in 'the wild'.  
  
Most often, you will need to find a data file (such as csv), and import it

## Importing data into R {.build}
For the purpose of this class, we have generated a **simulated** dataset of `r dataframe_join_about` to accompany the `r dataframe_name` dataset.  

To import the `r dataframe_join_file_name` file into RStudio, run the following:  
  
``r dataframe_join_name` <- read_csv(here::here("data", "`r dataframe_join_file_name`"))`

Go to code/  
Open 08_import_and_join.Rmd  
Complete the exercise to import this new dataset.  

## Importing data into R {.build}

You can also use the `readr` package to import data from a URL

For example, to load a dataset from a URL, run the following

```{r import_url, echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE}
path <- here::here("data", "phx.csv")
url <- "https://raw.githubusercontent.com/matthewhirschey/tidybiology-plusds/master/data/phx.csv"
patient_hx <- read_csv(url)

```
