---
title: "(Re)Introducing %>%"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## (Re)Introducing `%>%` {.build}

The `%>%` operator is a way of "chaining" together strings of commands that make reading your code easy.  
The following code chunk illustrates how `%>%` works:  
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  

```{r %>%, echo = FALSE, warning = FALSE}
df_input %>%
  select(colnames(df_character_select), colnames(df_numeric_select)) %>% 
  filter(!!rlang::sym(colnames(df_character_select)) == as.character(df_character_filter)) %>% 
  head()
```

## (Re)Introducing `%>%` {.build}

The previous code chunk does the following - it takes your dataset and then "pipes" it into `select()`, and then applies a `filter()` to the data.
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  
  
*the `head` function lists only the top n results -- convenient for long variables*
  
  
**When you see `%>%`, think "and then"**

## (Re)Introducing `%>%` {.build}

The alternative to using `%>%` is running the following code

filter(select(`r dataframe_name`, `r colnames(df_character_select)`, `r colnames(df_numeric_select)`), `r colnames(df_character_select)` == "`r df_character_filter`")

Although this is only one line as opposed to three, it's both more difficult to write and more difficult to read
