---
title: "ggplot: putting it all together"
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
#Load datasets we will use for the lecture
source(here::here("code", "healthcare.R")) #loads dataset-specific variables
```

# Lets take a glimpse at our dataset called `heart`
```{r}
glimpse(heart)
```

# Basics of a ggplot code
Below is an example of the most basic form of the ggplot code
```{r, eval=FALSE}
ggplot(data)+ #this will generate a blank plot for your data. At this point, ggplot knows what to plot, but not how.
  geom(mapping=aes(x, y)) #adding a geom tells ggplot how you would like to map your data
```

Use this template to make a simple ggplot.  
  
1. Use the `heart` dataset, 
2. Make a scatterplot comparing age vs. maximum heart rate
Recall that a "data dictionary" is located here: https://github.com/matthewhirschey/tidybiology-plusds

```{r first_plot}
ggplot(heart) +
  geom_point(aes(x = age, y = thalach))
```


Next, fit a smooth line (linear regression) to these points
*by adding another `geom`
```{r}
ggplot(heart, aes(x = age, y = thalach)) +
  geom_point() +
  geom_smooth(method = "lm")
```
  
Next, color by whether the patient has exercise-induced angina
```{r}
ggplot(heart, aes(x = age, y = thalach, color = exang)) +
  geom_point() +
  geom_smooth(method = "lm")
```

The last plot probably has your color based on a continuous scale; now change color by sex (a discrete variable)
```{r}
ggplot(heart, aes(x = age, y = thalach, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Next, fix the labels to capitalize, and be more informative
```{r}
ggplot(heart, aes(x = age, y = thalach, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Patient Age", y = "HR Max during Exercise")
```

Next, give the plot a title, by calling `title = "title"` within `labs`
And give the plot a caption, by calling `caption = "Plot by YOURNAME"` within `labs`
```{r}
ggplot(heart, aes(x = age, y = thalach, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Patient Age", y = "HR Max during Exercise", title = "Relationship between age and Max HR during exercise", caption = "Plot by Matt")
```

Next, use `facet_grid` to split this graph based on whether the pateint has exercise induced angina or not. Make sure to wrap your variable in `factor()` to convert the numeric into a factor variable. 
```{r}
ggplot(heart, aes(x = age, y = thalach, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Patient Age", y = "HR Max during Exercise", title = "Relationship between age and Max HR during exercise", caption = "Plot by Matt") +
  facet_grid(cols = vars(factor(exang)))
```

Next, change the theme to `light`
```{r}
ggplot(heart, aes(x = age, y = thalach, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Patient Age", y = "HR Max during Exercise", title = "Relationship between age and Max HR during exercise", caption = "Plot by Matt") +
  facet_grid(cols = vars(factor(exang))) +
  theme_light()
```

Lastly, save the object as a png image using `ggsave`
```{r}
ggsave("ggplot_together.png")
```
