---
title: "stringR"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

#data import
Running this script will import the datasets you need
```{r}
source(here::here("code", "healthcare.R"))
```

#take a glimpse at the `heart_joined` dataset to look for the variables we'll be searching within
```{r}
glimpse(heart_joined)
```

#How many people having an `"auntie" or "aunt"` in their health history?  
```{r}
sum(str_count(heart_joined$family_history, "unt"))
```

#how many people have Mother/mother but not grandmothers in their family history?
```{r}
sum(str_count(heart_joined$family_history, "^(M|m)other"))
```

#Beginning with `heart_joined`, create a new column called `statin` that detects whether or not a patient is on a statin, "and then" filter the dataset to contain only this subgroup of people. Call this new dataframe `statin_subgroup` 
How many people are in this subgroup?
```{r}
statin_subgroup <- heart_joined %>% 
  mutate(statin = str_detect(medication_hx, "statin")) %>% 
  filter(statin == TRUE)

#391 people
```


