---
title: "Getting to know your data"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## (Re)Introducing `%>%` {.build}

The `%>%` operator is a way of "chaining" together strings of commands that make reading your code easy.  
The following code chunk illustrates how `%>%` works:  
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  

```{r %>%, echo = FALSE, warning = FALSE}
df_input %>%
  select(colnames(df_character_select), colnames(df_numeric_select)) %>% 
  filter(!!rlang::sym(colnames(df_character_select)) == as.character(df_character_filter)) %>% 
  head()
```

## (Re)Introducing `%>%` {.build}

The previous code chunk does the following - it takes your dataset and then "pipes" it into `select()`, and then applies a `filter()` to the data.
  
``r dataframe_name` %>%`  
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`   
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`   
    `head()`  
  
*the `head` function lists only the top n results -- convenient for long variables*
  
  
**When you see `%>%`, think "and then"**

## (Re)Introducing `%>%` {.build}

The alternative to using `%>%` is running the following code

filter(select(`r dataframe_name`, `r colnames(df_character_select)`, `r colnames(df_numeric_select)`), `r colnames(df_character_select)` == "`r df_character_filter`")

Although this is only one line as opposed to three, it's both more difficult to write and more difficult to read

# R <br>Manipulating your data
Any questions?

## Introducing `dplyr` {.build}
  
dplyr is a package that contains a suite of functions that allow you to easily manipulate a dataset  
  
Some of the things you can do are -  
  
- select rows and columns that match specific criteria  
  
- create new variables (columns)  
  
- obtain summary statistics on individual groups within your datsets  
  
The main verbs we will cover are `select()`, `filter()`, `arrange()`, `mutate()`, and `summarise()`. These all combine naturally with `group_by()` which allows you to perform any operation "by group"

## `select()` specific columns from your dataset {.build}
The most basic `select()` is one where you comma separate a list of columns you want included

For example, if you only want to select the `r colnames(df_character_select)` and `r colnames(df_numeric_select)` columns, run the following code chunk

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`, `r colnames(df_numeric_select)`) %>%`  
    `head()`
    
```{r select_exercise, exercise=TRUE}

```

## `select()` {.build}

If you want to select all columns *except* `r colnames(df_character_select)`, run the following

``r dataframe_name` %>%`  
    `select(-`r colnames(df_character_select)`) %>%`  
    `head(5)`

```{r selectnot_exercise, exercise=TRUE}

```

## `select()` {.build}

Finally, you can provide a range of columns to return two columns and everything in between. For example

``r dataframe_name` %>%`   
    `select(`r colnames(df_character_select)`:`r colnames(df_numeric_select)`) %>%`  
    `head(5)`
    
```{r selectrange_exercise, exercise=TRUE}

```

## `filter()` rows based on certain condition(s) {.build}

The `filter()` verb evalutes a logical statement, and if a row meets the condition of this statement (i.e. is true) then it gets chosen (or "filtered").  
  
  
**All other rows are discarded**  

## `filter()` {.build}

Filtering can be performed on **categorical** data

``r dataframe_name` %>%`     
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`     
    `head(3)`  

```{r filter_exercise, exercise=TRUE}

```

Note that `filter()` only applies to rows, and has no effect on columns

## `filter()` {.build}

Filtering can also be performed on **numeric** data

For example, if you wanted to choose `r colnames(df_numeric_select)` with a value greater than `r df_numeric_filter`, you would run the following. 

``r dataframe_name` %>%`     
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`     
    `head(3)`  

```{r filter_exercise2, exercise=TRUE}

```

## `filter()` {.build}

To filter on multiple conditions, you can write a sequence of `filter()` commands

``r dataframe_name` %>%`       
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`") %>%`  
    `filter(`r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`    
    `head(3)`  

```{r filter_exercise3, exercise=TRUE}

```

## `filter()` {.build}

To avoid writing multiple `filter()` commands, multiple logical statements can be put inside a single `filter()` command, separated by commas

``r dataframe_name` %>%`           
    `filter(`r colnames(df_character_select)` == "`r df_character_filter`",`  
           ` `r colnames(df_numeric_select)` > `r df_numeric_filter`) %>%`      
    `head(3)`  

```{r filter_exercise4, exercise=TRUE}

```

## `arrange()` sorts rows {.build}

The input for arrange is one or many columns, and `arrange()` sorts the rows in ascending order i.e. from smallest to largest

For example, to sort rows from smallest to largest `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`   
  `arrange(`r colnames(df_numeric_select)`) %>%`   
  `head(3)`  

```{r arrange_exercise, exercise=TRUE}

```

## `arrange()` {.build}

To reverse this order, use the `desc()` function within `arrange()`

``r dataframe_name` %>%`  
  `arrange(desc(`r colnames(df_numeric_select)`)) %>%`  
  `head(3)`  

```{r arrange_exercise2, exercise=TRUE}

```


## `mutate()` {.build}

The `mutate()` verb, unlike the ones covered so far, creates new variable(s) i.e. new column(s). For example

``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(1)`  
  
The code chunk above takes all the elements of the column `r colnames(df_numeric_select)`, evaluates the square root of each element, and populates a new column called `new_col` with these results
  
## Try `mutate()` to make a new column
  
``r dataframe_name` %>%`    
  `mutate(new_col = sqrt(`r colnames(df_numeric_select)`)) %>%`    
  `head(3)`  
  
```{r mutate_exercise, exercise=TRUE}

```


## `summarise()` and `summarize()` {.build}

`summarize()` is one of the key functions in `dplyr`. It produces a new dataframe that aggregates that values of a column based on a certain condition.  
  
For example, to calculate the mean `r colnames(df_numeric_select)`, run the following

``r dataframe_name` %>%`      
  `summarise(mean(`r colnames(df_numeric_select)`))` 

```{r summarize_exercise, exercise=TRUE}

```

## `group_by()` {.build}

`group_by()` and `summarize()` are a powerful combination of functions to summarize by groups  
  
``r dataframe_name` %>%`  
  `group_by(`r colnames(df_character_select)`) %>%`  
  `summarise(mean(`r colnames(df_numeric_select)`))`  
  
```{r groupby_exercise, exercise=TRUE}

```
  
*remember this pair!  

## Saving a new dataset as an object

If you'd like to save the output of your wrangling as an object, you will need to use the `<-` operator

``r dataframe_name`_new <- `r dataframe_name` %>%`  
            `group_by(`r colnames(df_character_select)`) %>%`  
            `summarise(mean(`r colnames(df_numeric_select)`))`  
  
*Assigning the object the same name (i.e. ``r dataframe_name``) will overwrite the object  
  
```{r df_object_exercise, exercise=TRUE}

```
  
## Saving a new dataset as a file

To save ``r dataframe_name`_new` as a new file (e.g. csv), run the following:  
  
`write_csv(`r dataframe_name`_new, "`r dataframe_name`_new.csv")`
  
  
# R <br>Data Science Workflow
Any questions?

## Visualizing our dataset
  
Go to code/  
Open 03_rmd_practice.Rmd  
Complete the activity.
