---
title: "Tidybiology Tiny Lectures: ggplot "
subtitle: "ggplot: geoms"
author: "Matthew Hirschey, Ph.D."
date: "April 2020"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
library(patchwork)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## How are these two plots similar?
```{r double_plot, echo = FALSE, message=FALSE, warning=FALSE}
plot1 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_point() 

plot2 <- df_input %>% 
  ggplot(aes(!!sym(df_numeric1_name), !!sym(df_numeric2_name))) +
  geom_smooth() 

plot1 + plot2
```

## The `geom` is different between these plots
`geom` is short for geometric object, which is the visual object used to represent the data  
    
`plot1 <- ggplot(`r dataframe_name`) +`  
  `geom_point(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   
  
`plot2 <-  ggplot(`r df_numeric1_name`, `r df_numeric2_name`) +`  
  `geom_smooth(aes(`r df_numeric1_name`, `r df_numeric2_name`))`   

## Choosing your (geom)itries
Different data types require different plot types.  
  
When plotting your data, it is often helpful to take a glimpse at the data you intend to plot to know what kinds of variables you will be working with

`glimpse(`r dataframe_name`)`
```{r glimpse_again, exercise=TRUE}

```

## Geoms 
So now that you know your variable types, how do you know what geoms to use??  
  
Use the following resources to match your data type to the appropriate `geoms`

----
<a href="https://rstudio.com/resources/cheatsheets" target="_blank">https://rstudio.com/resources/cheatsheets/</a><br>
<center><iframe src="https://rstudio.com/resources/cheatsheets"></iframe></center>

----
<a href="https://www.data-to-viz.com/" target="_blank">https://www.data-to-viz.com/</a><br>
<center><iframe src="https://www.data-to-viz.com/"></iframe></center>

## Change this scatterplot code to draw boxplots {.smaller} 
`ggplot(`r dataframe_name`) + geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  
```{r boxplot_activity, exercise = TRUE}

```
  
Use the cheatsheet. Try your best guess.  

## What will this code do?
  
`ggplot(`r dataframe_name`) +`   
`geom_boxplot(aes(x=`r df_char1_name`,y=`r df_numeric1_name`)) +`  
`geom_point(aes(x=`r df_char1_name`,y=`r df_numeric1_name`))`  

## Try including two `geoms` for yourself
```{r 2geom_activity, exercise = TRUE}

```
  
**Each new geom adds a new layer**  

