---
title: "Introduction to RegEx"
output: 
  ioslides_presentation:
    widescreen: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(knitr)
library(here)
library(rlang)
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align="center")
source(here::here("code", "bespoke.R")) #loads custom objects
```

## Using Regular Expressions
That solution worked in this case, but was not very elegant, and might not work for all cases (what if there was a 'great aunt' in the list?)  

Or here is a more specific case for this data set.  
  
How many patients have a father with a history of disease? But we don't want to include grandfathers in the results. 
  
We can use something called **Regular Expressions**, aka **Regex**, to solve this

## Using Regular Expressions {.build}
Think of regex as a separate language, with it's own code, syntax, and rules.  
  
Regex rules allow complex matching patterns for strings, to ensure matching *exactly* the content desired
  
It is far too complex to cover in its entirely here, but here is one specific example.  
  
**GOAL**: identify all of the patients that have a father with a history of disease, but excluding grandfathers in the results. 
  
## Regular Expression Example {.build}
<div class="red2"> Example </div>  
We want to start with recognizing `father`.  

But then we want to make sure that we capture both `Father` and `father`. To accept either case f in the first spot we add `(F|f)`, so now our regex looks like `(F|f)ather`  
  
Lastly, we want this pattern to appear at the beginning of the word, so we add the regex `^` symbol.  

Our completed regex looks like:  
  
`str_count(heart_joined$family_history, "^(F|f)ather")` 

```{r regex1, exercise=TRUE}


```
## Regex Exercise
Go to code/  
Open 09_stringr.Rmd  
Complete the exercise to count mothers.    

## Regex resources
- The stringr cheatsheet shown above is a great resource  
- Here the tidyverse website with a link to its vignette  
  
<center><iframe src="https://stringr.tidyverse.org/articles/regular-expressions.html"></iframe></center>  